### 2022-12-27 paper/paper.tex
paper started

\documentclass[10pt]{extarticle}
\usepackage[margin=1in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{amsmath}
\title{Improving the Performance of Low Bandwidth Distributed Data Parallelism}
\author{Christopher Rae \\ \begin{small} email: 110641877@ea.edin.sch.uk \end{small}}

\date{}

\newcommand{\textify}[1]{\text{\begin{scriptsize}#1\end{scriptsize}}}

\begin{document}
\maketitle

\begin{abstract}
There is no doubt that networking is a bottleneck when training neural networks across multiple nodes. Having a low bandwidth or a high delay in time it takes for data to be transferred between nodes can drastically increase training time of a neural network. It has become industry standard to use fast networking with bandwidths of up to 100Gbps this makes the requirements for training across multiple nodes very expensive. \\
2. Research methods\\
3. is viable
\end{abstract}

\section{Introduction}
Distributed data parallelism is a widely used technique for training artificial intelligence (AI) models, especially when working with large datasets. It involves dividing the data across multiple nodes and training multiple copies of the model in parallel, which can significantly reduce the time required for training. However, the efficiency of this technique can be affected by several factors, including the bandwidth of the communication channels between the nodes.

Low WiFi bandwidths can have a significant impact on the training time of AI models using distributed data parallelism. When the communication channels between nodes have low bandwidth, it can lead to slower data transfer and synchronization between the nodes, which can result in slower training times. In addition, low WiFi bandwidths may also lead to increased latency in the communication between nodes, further diminishing the efficiency of the training process.

In this paper, we aim to explore the relationship between low WiFi bandwidths and the training time of AI models using distributed data parallelism. We will discuss the various ways in which low WiFi bandwidths can impact the communication between nodes and the overall efficiency of the training process. We will also present potential solutions for mitigating the negative effects of low WiFi bandwidths on model training time.

To provide a better understanding of the role that WiFi bandwidth plays in the training of AI models using distributed data parallelism, we will conduct experiments using a variety of datasets and model architectures. We will compare the training time of models trained using distributed data parallelism in high and low WiFi bandwidth environments to evaluate the impact of low WiFi bandwidths on the training process.

Overall, our goal is to shed light on the importance of WiFi bandwidth in the training of AI models using distributed data parallelism and to identify strategies for optimizing the training process in low bandwidth environments. We hope that the findings of this research will be useful for practitioners seeking to improve the efficiency of their model training process and for researchers studying the optimization of distributed data parallelism for AI model training.

\section{Methodology}
\subsection*{Distributed Data Parallelism}
The goal of distributed data parallelism is to split the data across the GPUs and computers in the cluster, so that the training process can be completed faster by leveraging the additional computational resources.

Here is an overview of the process:

\begin{enumerate}
	\item First, the data and model are partitioned or "split" across the GPUs and computers in 			  the cluster. This is typically done by dividing the data into smaller chunks and 	  				   assigning each chunk to a GPU or computer in the cluster.
	\item Next, the training process begins. On each GPU or computer in the cluster, the model 				  processes the data assigned to it and computes the gradients of the loss function with 		  respect to the model parameters.
	\item The gradients computed on each GPU or computer are then averaged together, and the 				  model parameters are updated using this average gradient. This process is known as 				  "gradient averaging" or "gradient synchronization."
	\item The process is then repeated, with each GPU or computer processing its assigned data 				  and computing gradients, until the model has been trained.
\end{enumerate}

By training the model in this way, the computation is distributed across multiple GPUs and computers, which can significantly reduce the time it takes to train the model. However, it also adds complexity to the training process, as the gradients must be averaged across the GPUs and computers and the model parameters must be kept consistent across the cluster. There are multiple ways of preforming this gradient averaging depending on the architecture of your GPUS and computers. In this paper we work with a synchronous architecture(all gpus have roughly the same comute power) for simplicity, but there is also the argument of a centralised versus decentralised topology.

In a centralized topology, the process of gradient averaging or gradient synchronization in step 3 of the distributed data parallelism process is typically performed on a central server or "parameter server." This is done by sending the gradients computed on each GPU or computer in the cluster to the central server, where they are averaged together and the model parameters are updated.

In a decentralized topology, the process of gradient averaging or gradient synchronization is performed in a decentralized manner, without the use of a central server or parameter server. This can be done using a number of different techniques, such as peer-to-peer communication or gossip protocols.

We find in this paper ---INSERT LINK--- that decentralized topologys are effected less by low bandwidths, and that is why we implment a decentralized architecture in our papers. To be more specific, we use our own implementation of the Baidu Ring AllReduce algorithm to average our gradients.

\subsection*{Gradient Quantization}
Model gradients are typically stored using 32-bit or 64-bit floating point values, in order to reduce the amount of data sent over the network we quantize our gradients.

When gradients under go out implementaion of quantization thier physical size is halved for example a 32-bit float the gradient will become 16-bit. This will reduce the accuracy of the model but will theoritcally half the communication time.

\subsection*{Gradient Sparcification}
Gradient sparsification is a technique that involves identifying and removing "zero" or nearly zero gradients from the gradients being transmitted between GPUs or computers in a distributed data parallelism setting. The goal is to reduce the size of the gradients without significantly affecting their quality, so that they can be transmitted more efficiently over the network.

In order to perform gradient sparsification we have to decide on a threshold to determine whether or not a gradient is a zero gradient or not, if the gradient is less than the threshold then it it deemed a zero gradient. In our implementation of gradient sparcification we accumulate zero gradients locally and sum them with gradients from the next iteration.

\subsection*{Larger Batch Sizes}
Increasing the batch size in a distributed training setup can also improve training efficiency by reducing the frequency of the gradient averaging step. This is because each device will process a larger amount of data before the gradients are averaged and applied to the model weights. This can reduce the overhead associated with performing the gradient averaging step, and can allow the model to make more efficient use of the available computation resources.

However, it is important to carefully tune the batch size, as increasing it too much can negatively impact the convergence and accuracy of the model. This is because a larger batch size can lead to a more noisy gradient estimate, which can make it more difficult for the model to learn effectively.

\section{Results}
\section{Conclusion}

\end{document}

### 2022-12-29 paper/paper.tex
updates

@@
-\documentclass[10pt]{extarticle}
+\documentclass[11pt]{extarticle}
 \usepackage[margin=1in]{geometry}
 \usepackage[parfill]{parskip}
-\usepackage{amsmath}
+\usepackage{amsmath, amsfonts}
 \title{Improving the Performance of Low Bandwidth Distributed Data Parallelism}
@@
 However, it also adds complexity to the training process, as the gradients must be averaged across the GPUs and computers and the model parameters must be kept consistent across the cluster.
-There are multiple ways of preforming this gradient averaging depending on the architecture of your GPUS and computers.
-In this paper we work with a synchronous architecture(all gpus have roughly the same comute power) for simplicity, but there is also the argument of a centralised versus decentralised topology.
+There are multiple ways of preforming this gradient averaging depending on the architecture of your GPUs and computers.
+In this paper we work with a synchronous architecture(all GPUs have roughly the same compute power) for simplicity, but there is also the argument of a centralised versus decentralised topology.
 
@@
 
-We find in this paper ---INSERT LINK--- that decentralized topologys are effected less by low bandwidths, and that is why we implment a decentralized architecture in our papers.
+We find in this paper ---INSERT LINK--- that decentralized topology's are effected less by low bandwidths, and that is why we implement a decentralized architecture in our papers.
 To be more specific, we use our own implementation of the Baidu Ring AllReduce algorithm to average our gradients.
 
+\subsection*{SGD}
+
+The training data is denoted as $\{x_i\}^n_{i=1} \subset \mathbb{R}^d$, and the model parameter vector is donated as\\ $w \in \mathbb{R}^d$.
+The goal of training is to solve the following optimization problem
+$$
+\min_w f
+(w) = \frac{1}{n} \sum^n_{i=1}L(w,x_i)
+$$
+We typically use a an algorithm such as stochastic gradient descent(SGD) to solve the equation above.
+In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w)] = {\nabla} f(w)$.
+SGD written as:
+
+$$
+w_{t+1} = w_t - \eta g_t(w_t)
+$$
+
+Where $g_t(w_t)$ is an estimate for the actual gradient ${\nabla} f(w_t)$ and $\eta$ is the learning rate.
+For simplicity we'll let $g$ represent $g_t(w_t)$.
+The convergence of SGD is largely influenced by the variance of $g$.
+In most implementations of distributed parallelism it is $g$ which is averaged between all of the replicas.
+
 \subsection*{Gradient Quantization}
-Model gradients are typically stored using 32-bit or 64-bit floating point values, in order to reduce the amount of data sent over the network we quantize our gradients.
+Gradient quantization is a technique used in distributed data parallelism to reduce the amount of data that needs to be transmitted over channels during training.
+It works by reducing the precision of the gradients that are transmitted between worker nodes.
 
-When gradients under go out implementaion of quantization thier physical size is halved for example a 32-bit float the gradient will become 16-bit.
-This will reduce the accuracy of the model but will theoritcally half the communication time.
+A model parameter is usually stored in memory as a 32bit floating point.
+There are $7, 800, 000, 000,\\ 000, 000,000,000,000,000,000,000,000,000,000,000,000,000,000$(38zeros) unique float32 numbers, but that said numbers like $0.000100$, $0.000101$ and $0.000102$ are very close.
+We can approximate these to number to be equal, this is the core concept of quantization.
+When we calculate
 
-\subsection*{Gradient Sparcification}
+If you have a vector $a \in \mathbb{R}^d$ where $a \sim \mathcal{N}(0.1,0.55)$ where we know the $\max a$ and $\min a$.
+We can split $a$ into equal size bins and represent each bin with either a 8bit or 16bit integer.
+For example, suppose we have a set of 100 bins, each representing a range of gradient values.
+If the gradient value for a particular parameter falls within the range of the first bin, it will be represented as a 1; if it falls within the range of the second bin, it will be represented as a 2; and so on.
+Then once it has been sent to the second node the node can look up what values the bin represents and replaces it with the excepted value of that bin.
+This allows us to represent each gradient value using a single integer, rather than a floating-point value, which can significantly reduce the amount of memory required to store gradients.
+The difference between using 8bit or 16bit integer is the number of bins, which effect not only the accuracy of the data but how much data is sent over the network.
+If we use 16 bit integers there can be 65,536 bins where as with 8bit there is only 256 which is a significant reduction.
+
+As mentioned above $a \sim \mathcal{N}(0.1,0.55)$ this tells us that that most of the values will be located around $0.1$ which means it may be more representative of $a$ if the there are more bins located around the mean than $\max a$ and $\min a$ so bins near the mean cover a lower range of values meaning they are occur at a higher frequency, bins closer to the min and the max cover a larger range of values.
+This type of quantization is called non-linear quantization.
+
+Modern models can contain billions of trainable parameters, which translates to billions of gradients.
+If we represent all of those gradient as 256 different integers we loose a lot of accuracy so what we do is we split gradients into chunks of around 2000 values and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk.
+there are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core.
+If one of the values in $a$ is $7$ this is a clear outlier this will effect one of the bins causing 1 bin to be wasted because of single outlier.
+By splitting up the data into chunks only a hand full of chunks will be effected by outliers.
+This type of quantization is called block-wise quantization.
+
+\subsubsection*{Formula}
+
+\subsection*{Gradient Sparsification}
 Gradient sparsification is a technique that involves identifying and removing "zero" or nearly zero gradients from the gradients being transmitted between GPUs or computers in a distributed data parallelism setting.
@@
 In order to perform gradient sparsification we have to decide on a threshold to determine whether or not a gradient is a zero gradient or not, if the gradient is less than the threshold then it it deemed a zero gradient.
-In our implementation of gradient sparcification we accumulate zero gradients locally and sum them with gradients from the next iteration.
+In our implementation of gradient sparsification we accumulate zero gradients locally and sum them with gradients from the next iteration.
+
+\subsubsection*{Formula}
+We denote sparsified gradients as $S(g, Z_t)$.
+We have to remember that $g \in \mathbb{R}^d$ so we let $g_i$ be the $i$-th component of the vector ($g=[g_1,\ldots,g_d]$).
+$Z$ is a vector with the same shape a $g$, when the $w$ is initialized at the start of the training process $Z_i = 0$.
+The function $S$ has 2 outputs the first being the new representation of the gradient of the weights at $t$, $v$ and the second being the accumulated gradients $Z_{t+1}$
+
+Sparsification of the gradients is shown as follows: 
+$$
+v_{t,i} =
+\left\{
+	\begin{array}{cc}
+		0 & \quad g_i + Z_{t,i}<\alpha \\
+		g_i + Z_{t,i}	& \quad g_i + Z_{t,i} \geq \alpha
+	\end{array}
+\right\}
+$$
+$$
+Z_{t+1,i} =
+\left\{
+	\begin{array}{cc}
+		Z_{t,i} + g_i & \quad g_i + Z_{t,i}<\alpha \\
+		0	& \quad g_i + Z_{t,i} \geq \alpha
+	\end{array}
+\right\}
+$$
+
+where $\alpha$ is the the threshold
 

### 2022-12-29 paper/paper.tex
updates to paper

@@
-\documentclass[11pt]{extarticle}
+\documentclass[11pt]{article}
 \usepackage[margin=1in]{geometry}
 \usepackage[parfill]{parskip}
-\usepackage{amsmath, amsfonts}
+\usepackage{amsmath, amsfonts,tabularx}
 \title{Improving the Performance of Low Bandwidth Distributed Data Parallelism}
@@
 \maketitle
+\tableofcontents
+\newpage
 
@@
 \section{Methodology}
-\subsection*{Distributed Data Parallelism}
+\subsection{Distributed Data Parallelism}
 The goal of distributed data parallelism is to split the data across the GPUs and computers in the cluster, so that the training process can be completed faster by leveraging the additional computational resources.
@@
 
-\subsection*{SGD}
+\subsection{SGD}
 
@@
 $$
+Where $L$ is the loss function.
 We typically use a an algorithm such as stochastic gradient descent(SGD) to solve the equation above.
@@
 
-\subsection*{Gradient Quantization}
+\subsection{Gradient Quantization}
 Gradient quantization is a technique used in distributed data parallelism to reduce the amount of data that needs to be transmitted over channels during training.
@@
 
-\subsubsection*{Formula}
+\subsubsection{Formula}
 
-\subsection*{Gradient Sparsification}
+\subsection{Gradient Sparsification}
 Gradient sparsification is a technique that involves identifying and removing "zero" or nearly zero gradients from the gradients being transmitted between GPUs or computers in a distributed data parallelism setting.
@@
 
-\subsubsection*{Formula}
+\subsubsection{Formula}
 We denote sparsified gradients as $S(g, Z_t)$.
@@
 
-\subsection*{Larger Batch Sizes}
+\subsection{Larger Batch Sizes}
+In machine learning, the batch size refers to the number of samples that will be processed before the model's weights are updated.
+The batch size is a hyperparameter that can be chosen when training a model.
+When using a batch size the SGD algorithm becomes mini-batch SGD it looks something like this:
+
+$$
+g_t(w_t) = \frac{1}{B_k} \sum_{i=i}^{B_k} {\nabla} f(w_t)
+$$
+$$
+w_{t+1} = w_t - \eta g_t(w_t)
+$$
+
 Increasing the batch size in a distributed training setup can also improve training efficiency by reducing the frequency of the gradient averaging step.
@@
 
-However, it is important to carefully tune the batch size, as increasing it too much can negatively impact the convergence and accuracy of the model.
-This is because a larger batch size can lead to a more noisy gradient estimate, which can make it more difficult for the model to learn effectively.
+One negative effect of large batch sizes is that they can lead to a decrease in generalization performance.
+This occurs because large batch sizes can result in an increase in the variance of the stochastic gradient descent (SGD) updates, which can cause the model to overfit to the training data.
+This overfitting can manifest as a decrease in performance on held-out test data or a decrease in the ability of the model to generalize to unseen data.
+
+Another negative effect of large batch sizes is that they can lead to a decrease in the convergence rate of the model.
+This occurs because large batch sizes can result in a reduction in the magnitude of the SGD updates, which can slow down the convergence of the model.
+In some cases, this can result in the model failing to converge at all.
+
+Furthermore, using large batch sizes can lead to an increase in the amount of memory required to store and process the data.
+This can be a problem for systems with limited memory or for models with a large number of parameters, as the large batch sizes may not fit in memory.
+
+There are a few things we can do to recude the negative effects of a large batch size.
+
+\subsection{Tests}
+\subsubsection{Models}
+Within our test we use a range of diferrent models. The most important factor when choosing a model is to make sure we ge a good range of sizes between out models. These are the models we have tested:
+
+\begin{center}
+\begin{tabular}{| p{3cm} | p{3cm} | p{3cm} | p{3cm} |} 
+\hline
+Model & Parameters & Dataset & Task \\ \hline
+RESNET-50 & 25,583,592 & ImageNet-1k & MoCo v2 \\ \hline
+VGG16 &  138,357,544 & CIFAR-100 & None \\ \hline
+Inception v4 & 42,711,400 & ImageNet-1k & MoCo v2\\ \hline
+Transformer & 0 & Multiple & GLUE \\ \hline
+RoBERTa-Large & 0 & Multiple & GLUE \\ \hline
+\end{tabular}
+\end{center}
+\subsubsection{testing methods}
+We start
 

### 2023-01-01 paper/paper.tex
paper updatd

@@
 \usepackage{amsmath, amsfonts,tabularx}
-\title{Improving the Performance of Low Bandwidth Distributed Data Parallelism}
+\title{Improving the Performance of Data Parallelism \\Over a Low Bandwidth Distributed }
 \author{Christopher Rae \\ \begin{small} email: 110641877@ea.edin.sch.uk \end{small}}
@@
 We typically use a an algorithm such as stochastic gradient descent(SGD) to solve the equation above.
-In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w)] = {\nabla} f(w)$.
+In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w)] = {\nabla} L_t(w)$.
 SGD written as:
@@
 
-Where $g_t(w_t)$ is an estimate for the actual gradient ${\nabla} f(w_t)$ and $\eta$ is the learning rate.
+Where $g_t(w_t)$ is an estimate for the actual gradient ${\nabla} L_t(w_t)$ and $\eta$ is the learning rate.
 For simplicity we'll let $g$ represent $g_t(w_t)$.
@@
 
-where $\alpha$ is the the threshold
+where $\alpha$ is the the threshold.
 
@@
 In machine learning, the batch size refers to the number of samples that will be processed before the model's weights are updated.
-The batch size is a hyperparameter that can be chosen when training a model.
+The batch size is a hyper parameter that can be chosen when training a model.
 When using a batch size the SGD algorithm becomes mini-batch SGD it looks something like this:
@@
 $$
-g_t(w_t) = \frac{1}{B_k} \sum_{i=i}^{B_k} {\nabla} f(w_t)
+g = \frac{1}{B_k} \sum_{i=1}^{B_k} {\nabla} L(w_t)
 $$
 $$
-w_{t+1} = w_t - \eta g_t(w_t)
+w_{t+1} = w_t - \eta g
 $$
 
+Where $B_k$ is the batch size.
 Increasing the batch size in a distributed training setup can also improve training efficiency by reducing the frequency of the gradient averaging step.
@@
 
-There are a few things we can do to recude the negative effects of a large batch size.
+One way to counteract the second point is to increase the learning rate $\eta$ so that the magnitude of the SGD updates are increased this usually works for a few of epochs but after a while has a limited effect.
+We plan to use a dynamic batch size.
+At the start of training a batch size is determined and the test error is calculated from the models untrained parameters.
+After each epoch the the the test error is check once the test error is half of the original the batch size is then doubled, once the test error has quartered the batch size is doubled again batch.
 
-\subsection{Tests}
-\subsubsection{Models}
-Within our test we use a range of diferrent models. The most important factor when choosing a model is to make sure we ge a good range of sizes between out models. These are the models we have tested:
+Another way which we can reduce the ngative effect of a lrge batch size is to use a optimizer algorithm other than SGD such as Adam.
+Adam (Adaptive Moment Estimation) is an optimization algorithm.
+It is an extension of the stochastic gradient descent (SGD) algorithm and combines the benefits of SGD with the benefits of other adaptive optimization algorithms such as RMSprop.
+
+One way that Adam helps to reduce the negative effects of using a large batch size is by using an adaptive learning rate.
+Adam maintains an exponential moving average of the gradients and the squared gradients, which it uses to compute an adaptive learning rate for each parameter.
+This adaptive learning rate is then used to update the parameters of the model, rather than using a fixed learning rate as in SGD.
+
+By using an adaptive learning rate, Adam is able to make more aggressive updates to the model's parameters when the gradients are small and make smaller updates when the gradients are large.
+This can help the model converge more quickly and reduce the negative effects of using a large batch size.
+
+In addition to using an adaptive learning rate, Adam also incorporates momentum into the optimization process.
+This helps the model to continue making progress even when it encounters noisy or oscillating gradients although this doesn't have an effect on the low bandwidth.
+
+\subsubsection{Formula for the Dynamic Batch Size}
+
+Let $E_e$ be the test error function at epoch $e$ which takes the parameters $w_t$ which is the model parameters at the time step $t$ at the start of epoch $e$.
+Let $b$ be the initial batch size and where      $\beta$ is the maximum batch size.
+\newpage
+\begingroup
+\Large
+$$
+\varepsilon = \max \left(1, 2^{(\lfloor E_0(w_0,x_R)/E_e(w_t, x_R) \rfloor -1)} \right)
+$$
+$$
+B_k = \min \left(\beta,b\varepsilon\right)
+$$
+\endgroup
+
+These equation explain dynamic batch sizing.
+first the scaling factor $\varepsilon$ is calculated by 2 to the power of the initial train error divided by the train error at epoch $e$ floored and subtracted by 1, we make sure that value is greater than 1 as if for some reason $E_0(w_0,x_R) < E_e(w_t, x_R)$ the batch size would decrease which is not what we want.
+Then its just a case of multiplying $b$ by your scaling factor    $\varepsilon$ and check to see if its greater than $\beta$.
+Please not that using the output of a loss function in stead of the test error is sufficient.
+
+\subsection{Models}
+Within our test we use a range of different models.
+The most important factor when choosing a model is to make sure we get a good range of sizes between out models.
+These are the models we have tested:
 
@@
 RESNET-50 & 25,583,592 & ImageNet-1k & MoCo v2 \\ \hline
-VGG16 &  138,357,544 & CIFAR-100 & None \\ \hline
-Inception v4 & 42,711,400 & ImageNet-1k & MoCo v2\\ \hline
-Transformer & 0 & Multiple & GLUE \\ \hline
-RoBERTa-Large & 0 & Multiple & GLUE \\ \hline
+VGG16 &  138,357,544 & CIFAR-10 & None \\ \hline
+Inception v4 & 42,711,400 & CIFAR-100 & None\\ \hline
+%Transformer & 0 & Multiple & GLUE \\ \hline
+%RoBERTa-Large & 0 & Multiple & GLUE \\ \hline
 \end{tabular}
 \end{center}
-\subsubsection{testing methods}
-We start
 
 \section{Results}
+We start by calculating a theoretical minimum training time.
+
 \section{Conclusion}

### 2023-01-02 paper/paper.tex
paper update

@@
 \usepackage[parfill]{parskip}
-\usepackage{amsmath, amsfonts,tabularx}
+\usepackage{amsmath, amsfonts,tabularx,graphicx, subfigure}
 \title{Improving the Performance of Data Parallelism \\Over a Low Bandwidth Distributed }
@@
 \maketitle
-\tableofcontents
-\newpage
 
@@
 \end{abstract}
+\tableofcontents
+\newpage
 
@@
 For example, suppose we have a set of 100 bins, each representing a range of gradient values.
-If the gradient value for a particular parameter falls within the range of the first bin, it will be represented as a 1; if it falls within the range of the second bin, it will be represented as a 2; and so on.
+If the gradient of a particular parameter falls within the range of the first bin, it will be represented as a 1; if it falls within the range of the second bin, it will be represented as a 2; and so on.
 Then once it has been sent to the second node the node can look up what values the bin represents and replaces it with the excepted value of that bin.
@@
 \hline
-Model & Parameters & Dataset & Task \\ \hline
-RESNET-50 & 25,583,592 & ImageNet-1k & MoCo v2 \\ \hline
-VGG16 &  138,357,544 & CIFAR-10 & None \\ \hline
-Inception v4 & 42,711,400 & CIFAR-100 & None\\ \hline
+Model & No. Parameters &  Gradeint Physical Size (MB)\\ \hline
+RESNET-50 & 25,583,592 &  94.1 \\ \hline
+VGG16 &  138,357,544 &  428.2 \\ \hline
+Inception v4 & 42,608,450 &  170.2\\ \hline
 %Transformer & 0 & Multiple & GLUE \\ \hline
@@
 We start by calculating a theoretical minimum training time. 
+\begin{figure}[h]
+  \centering
+  \subfigure[\begin{small}The Mean Time to Train 1 Epoch on 1 P100 GPU \end{small}]{
+  \includegraphics[width=6cm]{models.png}
+  }
+  \subfigure[ The Theoretical Time to Send Gradients]{
+  \includegraphics[width=6cm]{gradientTimes.png}  
+  }
+  \caption{\begin{small} a \& b \end{small}}
+\end{figure}
+
+In an ideal world training speed would increase linearly with the number of GPUs you have training the model.
+In our test to determine the results of Figure 1 (a) we train each model with a $B_k$ (batch size) of 32 and on the cifar 100 dataset of which we use 50000 images for training, this means that there are just under 1600 iterations in a single epoch.
+Figure 1(b) is determined by the size of the model divided by the speed of the network.
+
+\subsection{8 GPUs Theoretical}
+In out result we plan to compare the theoretical with the applied.
+The data is evenly split between each node in this case 8 GPUs  each on separate nodes, translating to each node training with a sample of 6250 images.
+By increasing the number of nodes from 1 to 8 we reduces the number of communication steps per epoch(and iterations) to around 200 but there is also more data sent in each communication step as each node is sending its own gradients to 8 other nodes.
+
+If each node was to individually send its gradients to all the other node, there would be 56 ($N(N-1)$ where $N$ is the number of nodes) separate messages sent between nodes, if it was worst case scenario where you are training VGG16 on 100MBps you would have around 4 minutes of time waiting for the communication per epoch.
+This is cause by only 1 node sending data at a time due to the networks constraints but as we use a more efficient ring all reduce algorithm, we should have communication times similar to what is shown in Figure 1 (b)
+
+\subsubsection{Quantization}
+We plan to use 8bit quantization using both non-linear and block-wise techniques.
+non-linear quantization doesn't effect bandwidth it just helps to reduce variance.
+Block-wise allows us to decrease the gradients size by a factor of 4 but adds $\frac{p}{2048}\times 256$ float16 numbers on top of the in order to undo the quantization.
+
+\subsection{8 GPUS Applied}
+\begin{Huge}
+\begin{center}
+COULD NOT ACCESS A COMPUTER
+\end{center}
+\end{Huge}
 

### 2023-01-03 paper/paper.tex
paper updated

@@
 On each GPU or computer in the cluster, the model 				  processes the data assigned to it and computes the gradients of the loss function with 		  respect to the model parameters.
+	\newpage
 	\item The gradients computed on each GPU or computer are then averaged together, and the 				  model parameters are updated using this average gradient.
@@
 There are multiple ways of preforming this gradient averaging depending on the architecture of your GPUs and computers.
-In this paper we work with a synchronous architecture(all GPUs have roughly the same compute power) for simplicity, but there is also the argument of a centralised versus decentralised topology.
+In this paper we work with a synchronous architecture for simplicity and as all of our GPUs have roughly the same compute power.
+There is also the argument of a centralised versus a decentralised topology.
 
@@
 
-We find in this paper ---INSERT LINK--- that decentralized topology's are effected less by low bandwidths, and that is why we implement a decentralized architecture in our papers.
+We find in this paper ---INSERT LINK--- that decentralized topology's are effected less by low bandwidths, and that is why we implement a decentralized architecture in our paper.
 To be more specific, we use our own implementation of the Baidu Ring AllReduce algorithm to average our gradients.
@@
 Where $L$ is the loss function.
-We typically use a an algorithm such as stochastic gradient descent(SGD) to solve the equation above.
+We typically use an algorithm such as stochastic gradient descent(SGD) to solve the equation above.
 In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w)] = {\nabla} L_t(w)$.
-SGD written as:
+SGD can be written as:
 
@@
 A model parameter is usually stored in memory as a 32bit floating point.
-There are $7, 800, 000, 000,\\ 000, 000,000,000,000,000,000,000,000,000,000,000,000,000,000$(38zeros) unique float32 numbers, but that said numbers like $0.000100$, $0.000101$ and $0.000102$ are very close.
-We can approximate these to number to be equal, this is the core concept of quantization.
-When we calculate
+There are $7, 800, 000, 000,\\ 000, 000,000,000,000,000,000,000,000,000,000,000,000,000,000$(38zeros) unique 32bit floating points, but that said 32bit floating points like $0.000100$, $0.000101$ and $0.000102$ are very close.
+We can approximate these numbers to be equal, this is the core concept of quantization.
 
@@
 If the gradient of a particular parameter falls within the range of the first bin, it will be represented as a 1; if it falls within the range of the second bin, it will be represented as a 2; and so on.
-Then once it has been sent to the second node the node can look up what values the bin represents and replaces it with the excepted value of that bin.
-This allows us to represent each gradient value using a single integer, rather than a floating-point value, which can significantly reduce the amount of memory required to store gradients.
+Then once the quantized $a$ has been sent to another node that node can look up what values the bin represents and replaces all the gradients in the bin with that value.
+This allows us to represent each gradient value using a single integer, rather than a floating-point value, which can significantly reduce the amount of bandwidth required to when sending gradients.
 The difference between using 8bit or 16bit integer is the number of bins, which effect not only the accuracy of the data but how much data is sent over the network.
@@
 
+\subsubsection{Non-Linear Quantization}
+
 As mentioned above $a \sim \mathcal{N}(0.1,0.55)$ this tells us that that most of the values will be located around $0.1$ which means it may be more representative of $a$ if the there are more bins located around the mean than $\max a$ and $\min a$ so bins near the mean cover a lower range of values meaning they are occur at a higher frequency, bins closer to the min and the max cover a larger range of values.
-This type of quantization is called non-linear quantization.
+Instead of all the bins covering an equal range of values at a common interval each bin will contian an equal range of values(if you have 200 gradients and 10 bins each bin will contain 20 gradients).
+This value assigned to the the bin is the mean of the values in the bin.
+
+\subsubsection{Block-Wise Quantization}
 
 Modern models can contain billions of trainable parameters, which translates to billions of gradients.
-If we represent all of those gradient as 256 different integers we loose a lot of accuracy so what we do is we split gradients into chunks of around 2000 values and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk.
-there are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core.
-If one of the values in $a$ is $7$ this is a clear outlier this will effect one of the bins causing 1 bin to be wasted because of single outlier.
-By splitting up the data into chunks only a hand full of chunks will be effected by outliers.
-This type of quantization is called block-wise quantization.
-
-\subsubsection{Formula}
+If we represent all of those gradient as 256 different integers we loose a lot of accuracy and make it alot harder for the model to converge, so what we do is we split gradients into chunks of around 2000 or 4000 items(depending on the size of the model) and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk.
+there are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core makeing the quantixzation step alot faster.
+If one of the values in $a$ was $7$ it is a clear outlier, this will effect one of the bins causing 1 bin to be wasted because of single outlier.
+By splitting up the data into chunks only a one of chunks would be effected by this outliers.
 
@@
 We have to remember that $g \in \mathbb{R}^d$ so we let $g_i$ be the $i$-th component of the vector ($g=[g_1,\ldots,g_d]$).
-$Z$ is a vector with the same shape a $g$, when the $w$ is initialized at the start of the training process $Z_i = 0$.
-The function $S$ has 2 outputs the first being the new representation of the gradient of the weights at $t$, $v$ and the second being the accumulated gradients $Z_{t+1}$
+$Z \in \mathbb{R}^d$ is a vector with the same shape a $g$, when the $w$ is initialized at the start of the training process $Z_{0,i} = 0$.
+The function $S$ has 2 outputs the first being $v$ the new representation of the gradient of the weights at iteration $t$, and the second being the accumulated gradients $Z_{t+1}$
 
@@
 	\begin{array}{cc}
-		0 & \quad g_i + Z_{t,i}<\alpha \\
-		g_i + Z_{t,i}	& \quad g_i + Z_{t,i} \geq \alpha
+		0 & \quad \left| g_i + Z_{t,i}\right| < \alpha \\
+		g_i + Z_{t,i} & \quad \left| g_i + Z_{t,i}\right| \geq \alpha
 	\end{array}
@@
 	\begin{array}{cc}
-		Z_{t,i} + g_i & \quad g_i + Z_{t,i}<\alpha \\
-		0	& \quad g_i + Z_{t,i} \geq \alpha
+		Z_{t,i} + g_i & \quad \left| g_i + Z_{t,i}\right| < \alpha \\
+		0	& \quad \left| g_i + Z_{t,i}\right| \geq \alpha
 	\end{array}
@@
 We plan to use a dynamic batch size.
-At the start of training a batch size is determined and the test error is calculated from the models untrained parameters.
-After each epoch the the the test error is check once the test error is half of the original the batch size is then doubled, once the test error has quartered the batch size is doubled again batch.
+At the start of training a an inital batch size is determined, and the primary test error is calculated from the models untrained parameters.
+After each epoch the the the test error is checked, once the test error is half of the primary test error the batch size gets doubled, once the test error is a quarter of the primary teset error the batch size is doubled again batch.
 
-Another way which we can reduce the ngative effect of a lrge batch size is to use a optimizer algorithm other than SGD such as Adam.
+Another way which we can reduce the negative effect of a large batch size is to use a differnet optimizer algorithm to SGD such as Adam.
 Adam (Adaptive Moment Estimation) is an optimization algorithm.
@@
 Let $b$ be the initial batch size and where      $\beta$ is the maximum batch size.
-\newpage
+
 \begingroup
@@
 These equation explain dynamic batch sizing.
-first the scaling factor $\varepsilon$ is calculated by 2 to the power of the initial train error divided by the train error at epoch $e$ floored and subtracted by 1, we make sure that value is greater than 1 as if for some reason $E_0(w_0,x_R) < E_e(w_t, x_R)$ the batch size would decrease which is not what we want.
-Then its just a case of multiplying $b$ by your scaling factor    $\varepsilon$ and check to see if its greater than $\beta$.
+First the scaling factor $\varepsilon$ where $E_0(w_0,x_R)$ is the primary test error, we make sure that value of $\varepsilon$ is greater than 1 as if for some reason $E_0(w_0,x_R) < E_e(w_t, x_R)$ the batch size would decrease which is not what we want.
+Then its just a case of multiplying $b$ by your scaling factor $\varepsilon$ and checking to see if its greater than $\beta$.
 Please not that using the output of a loss function in stead of the test error is sufficient.
@@
 \subsection{8 GPUs Theoretical}
-In out result we plan to compare the theoretical with the applied.
-The data is evenly split between each node in this case 8 GPUs  each on separate nodes, translating to each node training with a sample of 6250 images.
+In out result we plan to compare the theoretical with the applied.In this case 8 GPUs were each node has 1 GPU, the data is evenly split between each node translating to each node training with a sample of 6250 images.
 By increasing the number of nodes from 1 to 8 we reduces the number of communication steps per epoch(and iterations) to around 200 but there is also more data sent in each communication step as each node is sending its own gradients to 8 other nodes.
 
-If each node was to individually send its gradients to all the other node, there would be 56 ($N(N-1)$ where $N$ is the number of nodes) separate messages sent between nodes, if it was worst case scenario where you are training VGG16 on 100MBps you would have around 4 minutes of time waiting for the communication per epoch.
-This is cause by only 1 node sending data at a time due to the networks constraints but as we use a more efficient ring all reduce algorithm, we should have communication times similar to what is shown in Figure 1 (b)
+If each node was to individually send its gradients to all the other node, there would be 56 ($N(N-1)$ where $N$ is the number of nodes) separate messages sent between nodes, if we were dealing with the worst case scenario where you are training VGG16 on 100MBps you would have around 4 minutes of communication per epoch.
+This is cause by only 1 node sending data at a time while the rest of the nodes are idol caused by the networks constraints.
+As we are using a more efficient ring all reduce algorithm, we should have communication times similar to what is shown in Figure 1 (b)
 
@@
 non-linear quantization doesn't effect bandwidth it just helps to reduce variance.
-Block-wise allows us to decrease the gradients size by a factor of 4 but adds $\frac{p}{2048}\times 256$ float16 numbers on top of the in order to undo the quantization.
+Block-wise allows us to decrease the gradients size by a factor of 4 but adds $\frac{p}{k}\times 256$ float16 numbers on top of the quatized gradients in order to undo the quantization where $p$ is the number of parameters and $k$ is the value that determines the number of block usually 2048 or 4096.
+For the example of vgg16 you decrease the graidents to around 110 MB in size but add an extra size of 8,000,000 32 bit floating points.
+
+\subsection{Quatization}
 

### 2023-01-04 paper/paper.tex
update paper

@@
 \usepackage{amsmath, amsfonts,tabularx,graphicx, subfigure}
-\title{Improving the Performance of Data Parallelism \\Over a Low Bandwidth Distributed }
-\author{Christopher Rae \\ \begin{small} email: 110641877@ea.edin.sch.uk \end{small}}
+\title{Improving the Performance of Distributed Data Parallelism \\Over a Low Bandwidth}
+\author{Christopher Rae \\ \begin{small} email: raecd123@gmail.com \end{small}}
 
@@
   \centering
-  \subfigure[\begin{small}The Mean Time to Train 1 Epoch on 1 P100 GPU \end{small}]{
+  \subfigure[\begin{small}Training 1 Epoch on 1 P100 GPU \end{small}]{
   \includegraphics[width=6cm]{models.png}
@@
 
-\subsubsection{Quantization}
 We plan to use 8bit quantization using both non-linear and block-wise techniques.
@@
 
-\subsection{Quatization}
+When using sparsification with quantization, sparsification occours before quantization and quantizaion only occours on the non zero grads.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=7cm]{zeroGrads.png}
+	\caption{\begin{small} Percentage of Zeros Grads in the First Epoch\end{small}}
+\end{figure}
+
+We see that as the model gets closer to converging the number of zero grads increases, the graph also shows us that when we have a threshold of 0.00001 we drop over 80\% of our gradients this seems like it would have a large effect on the training process this is something we will address in our applied section of the results.
+If we drop 80\% that would also mean you are only sending 20\% of gradients being sent between nodes over the network, on a network with a low bandwidth.
+
+By doubling the batch size you should half the communication time as there half as many communication step taking place in an epoch.
+Whilest using a dynamic batch size it can take quite a long time before the batch size doubles as the it takes multiple epoch before the train error halfs.
 

### 2023-01-04 paper/paper.tex
paper update

@@
 \begin{abstract}
-There is no doubt that networking is a bottleneck when training neural networks across multiple nodes.
-Having a low bandwidth or a high delay in time it takes for data to be transferred between nodes can drastically increase training time of a neural network.
-It has become industry standard to use fast networking with bandwidths of up to 100Gbps this makes the requirements for training across multiple nodes very expensive.
-\\
-2. Research methods\\
-3. is viable
+There is no doubt that networking is a bottleneck when training neural networks across multiple GPUs and computers.
+Having a low bandwidth or a high latency can drastically increase training time of a neural network.
+In this paper we look at a variety of methods to reduce the effect of a low bandwidth including gradient sparsification, gradient quantization and a using a large batch size.
+We find ...
+COULDN'T FIND COMPUTER
 \end{abstract}
@@
 non-linear quantization doesn't effect bandwidth it just helps to reduce variance.
-Block-wise allows us to decrease the gradients size by a factor of 4 but adds $\frac{p}{k}\times 256$ float16 numbers on top of the quatized gradients in order to undo the quantization where $p$ is the number of parameters and $k$ is the value that determines the number of block usually 2048 or 4096.
-For the example of vgg16 you decrease the graidents to around 110 MB in size but add an extra size of 8,000,000 32 bit floating points.
+Block-wise allows us to decrease the gradients size by a factor of 4 but adds $\frac{p}{k}\times 256$ float16 numbers on top of the quantized gradients in order to undo the quantization where $p$ is the number of parameters and $k$ is the value that determines the number of block usually 2048 or 4096.
+For the example of vgg16 you decrease the gradients to around 110 MB in size but add an extra size of 8,000,000 32 bit floating points.
 
-When using sparsification with quantization, sparsification occours before quantization and quantizaion only occours on the non zero grads.
+When using sparsification with quantization, sparsification occurs before quantization and quantization only occurs on the non zero grads.
 
@@
 By doubling the batch size you should half the communication time as there half as many communication step taking place in an epoch.
-Whilest using a dynamic batch size it can take quite a long time before the batch size doubles as the it takes multiple epoch before the train error halfs.
+Whilst using a dynamic batch size it can take quite a long time before the batch size doubles as the it takes multiple epoch before the train error halves.
 

### 2023-02-17 paper/paper.tex
remove section on adam

@@
 As mentioned above $a \sim \mathcal{N}(0.1,0.55)$ this tells us that that most of the values will be located around $0.1$ which means it may be more representative of $a$ if the there are more bins located around the mean than $\max a$ and $\min a$ so bins near the mean cover a lower range of values meaning they are occur at a higher frequency, bins closer to the min and the max cover a larger range of values.
-Instead of all the bins covering an equal range of values at a common interval each bin will contian an equal range of values(if you have 200 gradients and 10 bins each bin will contain 20 gradients).
+Instead of all the bins covering an equal range of values at a common interval each bin will contain an equal range of values(if you have 200 gradients and 10 bins each bin will contain 20 gradients).
 This value assigned to the the bin is the mean of the values in the bin.
@@
 If we represent all of those gradient as 256 different integers we loose a lot of accuracy and make it alot harder for the model to converge, so what we do is we split gradients into chunks of around 2000 or 4000 items(depending on the size of the model) and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk.
-there are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core makeing the quantixzation step alot faster.
+there are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core makeing the quantization step a lot faster.
 If one of the values in $a$ was $7$ it is a clear outlier, this will effect one of the bins causing 1 bin to be wasted because of single outlier.
@@
 We plan to use a dynamic batch size.
-At the start of training a an inital batch size is determined, and the primary test error is calculated from the models untrained parameters.
-After each epoch the the the test error is checked, once the test error is half of the primary test error the batch size gets doubled, once the test error is a quarter of the primary teset error the batch size is doubled again batch.
+At the start of training a an initial batch size is determined, and the primary test error is calculated from the models untrained parameters.
+After each epoch the the the test error is checked, once the test error is half of the primary test error the batch size gets doubled, once the test error is a quarter of the primary test error the batch size is doubled again batch.
 
-Another way which we can reduce the negative effect of a large batch size is to use a differnet optimizer algorithm to SGD such as Adam.
+Another way which we can reduce the negative effect of a large batch size is to use a different optimizer algorithm to SGD A common choice is Adam (Adaptive Moment Estimation) but adam has been found to scale poorly with large batch sizes.
+There are 2 optimizers that scale well with large batch sizes, LARS (Layer-wise Adaptive Rate Scaling) \& LAMB (Layer-wise Adaptive Moments for Batch Training)
+
+\iffalse
+such as Adam.
 Adam (Adaptive Moment Estimation) is an optimization algorithm.
@@
 This helps the model to continue making progress even when it encounters noisy or oscillating gradients although this doesn't have an effect on the low bandwidth.
+\fi
 
@@
 \hline
-Model & No. Parameters &  Gradeint Physical Size (MB)\\ \hline
+Model & No. Parameters &  Gradient Physical Size (MB)\\ \hline
 RESNET-50 & 25,583,592 &  94.1 \\ \hline

### 2023-02-17 paper/paper.tex
updates to conclusion
finished LAMB section and started bib

@@
 \usepackage[parfill]{parskip}
-\usepackage{amsmath, amsfonts,tabularx,graphicx, subfigure}
-\title{Improving the Performance of Distributed Data Parallelism \\Over a Low Bandwidth}
+\usepackage{amsmath, amsfonts,tabularx,graphicx, subfigure, mathtools}
+\title{Improving the Performance of Distributed Data Parallelism \\Over Low Communication  Bandwidths}
 \author{Christopher Rae \\ \begin{small} email: raecd123@gmail.com \end{small}}
@@
 
-\newcommand{\textify}[1]{\text{\begin{scriptsize}#1\end{scriptsize}}}
+%\newcommand{\textify}[1]{\text{\begin{scriptsize}#1\end{scriptsize}}}
+\DeclarePairedDelimiter\norm{\lVert}{\rVert}
 
@@
 
-We find in this paper ---INSERT LINK--- that decentralized topology's are effected less by low bandwidths, and that is why we implement a decentralized architecture in our paper.
+We find in this paper \cite{lian2017can} that decentralized topology's are effected less by low bandwidths, and that is why we implement a decentralized architecture in our paper.
 To be more specific, we use our own implementation of the Baidu Ring AllReduce algorithm to average our gradients.
@@
 We typically use an algorithm such as stochastic gradient descent(SGD) to solve the equation above.
-In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w)] = {\nabla} L_t(w)$.
+In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w_{t-1})] = {\nabla} L(w_{t-1})$.
 SGD can be written as:
@@
 $$
-w_{t+1} = w_t - \eta g_t(w_t)
+w_{t} = w_{t-1} - \eta g_t(w_{t-1})
 $$
 
-Where $g_t(w_t)$ is an estimate for the actual gradient ${\nabla} L_t(w_t)$ and $\eta$ is the learning rate.
+Where $g_t(w_{t-1})$ is an estimate for the actual gradient ${\nabla} L_t(w_t)$ and $\eta$ is the learning rate.
 For simplicity we'll let $g$ represent $g_t(w_t)$.
@@
 	\begin{array}{cc}
-		0 & \quad \left| g_i + Z_{t,i}\right| < \alpha \\
-		g_i + Z_{t,i} & \quad \left| g_i + Z_{t,i}\right| \geq \alpha
+		0 & \quad \left| g_i + Z_{t-1,i}\right| < \alpha \\
+		g_i + Z_{t-1,i} & \quad \left| g_i + Z_{t-1,i}\right| \geq \alpha
 	\end{array}
@@
 $$
-Z_{t+1,i} =
+Z_{t,i} =
 \left\{
 	\begin{array}{cc}
-		Z_{t,i} + g_i & \quad \left| g_i + Z_{t,i}\right| < \alpha \\
-		0	& \quad \left| g_i + Z_{t,i}\right| \geq \alpha
+		Z_{t-1,i} + g_i & \quad \left| g_i + Z_{t-1,i}\right| < \alpha \\
+		0	& \quad \left| g_i + Z_{t-1,i}\right| \geq \alpha
 	\end{array}
@@
 \subsection{Larger Batch Sizes}
-In machine learning, the batch size refers to the number of samples that will be processed before the model's weights are updated.
+In machine learning, we have the concept of batch sizes this can be thought of as the number of samples that will be processed before the model's weights are updated.
 The batch size is a hyper parameter that can be chosen when training a model.
@@
 $$
-g = \frac{1}{B_k} \sum_{i=1}^{B_k} {\nabla} L(w_t)
+g_{t-1} = \frac{1}{B_k} \sum_{i=1}^{B_k} {\nabla} L(w_{t-1})
 $$
 $$
-w_{t+1} = w_t - \eta g
+w_{t} = w_{t-1} - \eta g_{t}
 $$
@@
 One way to counteract the second point is to increase the learning rate $\eta$ so that the magnitude of the SGD updates are increased this usually works for a few of epochs but after a while has a limited effect.
-We plan to use a dynamic batch size.
+We plan to use a adaptive batch size \cite{devarakonda2017adabatch}.
 At the start of training a an initial batch size is determined, and the primary test error is calculated from the models untrained parameters.
@@
 
-Another way which we can reduce the negative effect of a large batch size is to use a different optimizer algorithm to SGD A common choice is Adam (Adaptive Moment Estimation) but adam has been found to scale poorly with large batch sizes.
-There are 2 optimizers that scale well with large batch sizes, LARS (Layer-wise Adaptive Rate Scaling) \& LAMB (Layer-wise Adaptive Moments for Batch Training)
+Another way which we can reduce the negative effect of a large batch size is to use a different optimizer algorithm to SGD, a common alternative is Adam (Adaptive Moment Estimation) the problem with ADAM is that ADAM has been found to scale poorly with large batch sizes.
+There are 2 optimizers that scale well with large batch sizes, LARS (Layer-wise Adaptive Rate Scaling) \cite{you2017scaling} \& LAMB (Layer-wise Adaptive Moments for Batch Training) \cite{you2019large}.
+We have chosen to use LAMB as our optimiser.
 
-\iffalse
-such as Adam.
-Adam (Adaptive Moment Estimation) is an optimization algorithm.
-It is an extension of the stochastic gradient descent (SGD) algorithm and combines the benefits of SGD with the benefits of other adaptive optimization algorithms such as RMSprop.
+LAMB is pretty similar to ADAM.
+We start by computing the gradients $g_t$ at time step $t$, then much like ADAM we calculate the first $m$ and second $v$ moment of each layer $l$ using the following equations:
+\begin{equation*} \label{1mom}
+m^l_t = \beta _1 m^l_{t-1} + (1-\beta _1) g^l_t
+\end{equation*}
+\begin{equation*} \label{2mom}
+v^l_t = \beta _2 v^l_{t-1} + (1-\beta _2) g^l_t \odot g^l_t
+\end{equation*}
+Then we bias correct for both of the moments:
+\begin{equation*} \label{b1}
+\hat{m}^l_t = m^l_t /(1-\beta ^t_1)
+\end{equation*}
+\begin{equation*} \label{b2}
+\hat{v}^l_t = v^l_t /(1-\beta ^t_2)
+\end{equation*}
 
-One way that Adam helps to reduce the negative effects of using a large batch size is by using an adaptive learning rate.
-Adam maintains an exponential moving average of the gradients and the squared gradients, which it uses to compute an adaptive learning rate for each parameter.
-This adaptive learning rate is then used to update the parameters of the model, rather than using a fixed learning rate as in SGD.
+In these equations $\beta _{1/2}$ are the coefficient of the first/second moment. After we calculate the moments we trust ratio:
 
-By using an adaptive learning rate, Adam is able to make more aggressive updates to the model's parameters when the gradients are small and make smaller updates when the gradients are large.
-This can help the model converge more quickly and reduce the negative effects of using a large batch size.
+\begin{equation*}
+r^l_t = 1.0
+\end{equation*}
+\begin{equation*}
+r_1 = \phi(\norm{w^l_{t-1}})
+\end{equation*}
+$\phi(\parallel x \parallel)$ where $x \in \mathbb{R}^d$ is the same as doing $\sqrt{\sum^d_{i=1} x_i^2}$
+\begin{equation*}\label{r2}
+r_2= \norm[\bigg]{\frac{\hat{m}^l_t}{\sqrt{\hat{v}^l_t + \epsilon}} + \lambda w^l_{t-1}}
+\end{equation*}
+\begin{equation*}
+r^l_t =
+\left\{
+	\begin{array}{cc}
+		r_1/r_2 & \quad r_1 > 0 \text{ and } r_2 > 0 \\
+		1 & \quad r_1 < 0 \text{ or } r_" < 0
+	\end{array}
+\right\}
+\end{equation*}
+$r_1$ is the norm of the gradients and $r_2$ is the element-wise weight decay.
+To make sure that our value is $r^l_t$ isn’t too large or too small we have and upper and lower boundary $a$ and $b$.
+$\epsilon$ in equation \ref{r2} is to prevent a division by 0 error.
 
-In addition to using an adaptive learning rate, Adam also incorporates momentum into the optimization process.
-This helps the model to continue making progress even when it encounters noisy or oscillating gradients although this doesn't have an effect on the low bandwidth.
-\fi
+our final step is to upgrade the parameters:
+\begin{equation*}
+w^l_t = w^l_{t-1} - \eta r^l_t  \left(\frac{\hat{m}^l_t}{\sqrt{\hat{v}^l_t + \epsilon}} 
++ \lambda w^l_{t-1}\right)
+\end{equation*}
 
-\subsubsection{Formula for the Dynamic Batch Size}
+\subsubsection{Formula for the Adaptive Batch Size}
 
@@
 
-These equation explain dynamic batch sizing.
+These equation explain adaptive batch sizing.
 First the scaling factor $\varepsilon$ where $E_0(w_0,x_R)$ is the primary test error, we make sure that value of $\varepsilon$ is greater than 1 as if for some reason $E_0(w_0,x_R) < E_e(w_t, x_R)$ the batch size would decrease which is not what we want.
@@
 Within our test we use a range of different models.
-The most important factor when choosing a model is to make sure we get a good range of sizes between out models.
-These are the models we have tested:
+The most important factor when choosing a model is to make sure we get a good range of sizes between our models.
+These are the models we train to get results:
 
@@
 Inception v4 & 42,608,450 &  170.2\\ \hline
+%BERT & 109,786,368 & 
 %Transformer & 0 & Multiple & GLUE \\ \hline
@@
 
+\newpage
 \section{Results}
-We start by calculating a theoretical minimum training time. 
 \begin{figure}[h]
@@
 
-\subsection{8 GPUs Theoretical}
+\subsection{Baseline}
 In out result we plan to compare the theoretical with the applied.In this case 8 GPUs were each node has 1 GPU, the data is evenly split between each node translating to each node training with a sample of 6250 images.
-By increasing the number of nodes from 1 to 8 we reduces the number of communication steps per epoch(and iterations) to around 200 but there is also more data sent in each communication step as each node is sending its own gradients to 8 other nodes.
+By increasing the number of nodes from 1 to 8 we reduces the number of communication steps per epoch to around 200 but there is also more data sent in each communication step as each node is sending its own gradients to 8 other nodes.
 
-If each node was to individually send its gradients to all the other node, there would be 56 ($N(N-1)$ where $N$ is the number of nodes) separate messages sent between nodes, if we were dealing with the worst case scenario where you are training VGG16 on 100MBps you would have around 4 minutes of communication per epoch.
+If each node was to individually send its gradients to all the other node, there would be $N(N-1)$ messages being sent between nodes, where $N$ is the number of nodes, in the case of 8 nodes that is $56$ messages.
+If we were dealing with our worst case scenario where you are training VGG16 on 100MBps you could take multiple hours ($4*200*56$ seconds) per epoch.
 This is cause by only 1 node sending data at a time while the rest of the nodes are idol caused by the networks constraints.
-As we are using a more efficient ring all reduce algorithm, we should have communication times similar to what is shown in Figure 1 (b)
+As we are using a more efficient ring all reduce algorithm, we should have a much lower communication time than multiple hours.
 
@@
 Block-wise allows us to decrease the gradients size by a factor of 4 but adds $\frac{p}{k}\times 256$ float16 numbers on top of the quantized gradients in order to undo the quantization where $p$ is the number of parameters and $k$ is the value that determines the number of block usually 2048 or 4096.
-For the example of vgg16 you decrease the gradients to around 110 MB in size but add an extra size of 8,000,000 32 bit floating points.
+For the example of vgg16 you decrease the gradients to around 110 MB in size but add the additional size of an extra size of 8,000,000 32 bit floating points.
 
@@
 
-We see that as the model gets closer to converging the number of zero grads increases, the graph also shows us that when we have a threshold of 0.00001 we drop over 80\% of our gradients this seems like it would have a large effect on the training process this is something we will address in our applied section of the results.
-If we drop 80\% that would also mean you are only sending 20\% of gradients being sent between nodes over the network, on a network with a low bandwidth.
+We see that as the model gets closer to converging the number of zero grads increases, the graph also shows us that when we have a threshold of 0.00001 we drop over 80\% of our gradients this seems like it would have a large effect on the training process this is something we will address in our Benchmarks.
+If we drop 80\%, you are only sending 20\% of gradients to other  nodes.
 
 By doubling the batch size you should half the communication time as there half as many communication step taking place in an epoch.
-Whilst using a dynamic batch size it can take quite a long time before the batch size doubles as the it takes multiple epoch before the train error halves.
+Whilst using a adaptive batch size it can take quite a long time before the batch size doubles as the it takes multiple epoch before the train error halves.
 
-\subsection{8 GPUS Applied}
+\subsection{Benchmarks}
 \begin{Huge}
@@
 \section{Conclusion}
+We were not able to access a computer so were not able to validate whether the techniques we implemented had a beneficial effect on the training time.
+
+\bibliographystyle{plain}
+\bibliography{paperRef}
 

### 2023-02-19 paper/paper.tex
updates to paper

@@
 \section{Introduction}
-Distributed data parallelism is a widely used technique for training artificial intelligence (AI) models, especially when working with large datasets.
+Distributed data parallelism is a widely used technique for training Artificial Intelligence (AI) models, especially when working with large datasets.
 It involves dividing the data across multiple nodes and training multiple copies of the model in parallel, which can significantly reduce the time required for training.
@@
 Low WiFi bandwidths can have a significant impact on the training time of AI models using distributed data parallelism.
-When the communication channels between nodes have low bandwidth, it can lead to slower data transfer and synchronization between the nodes, which can result in slower training times.
+When the communication channels between nodes have low bandwidth, this can lead to slower data transfer and synchronization between the nodes, which can in turn result in slower training times.
 In addition, low WiFi bandwidths may also lead to increased latency in the communication between nodes, further diminishing the efficiency of the training process.
@@
 There are multiple ways of preforming this gradient averaging depending on the architecture of your GPUs and computers.
-In this paper we work with a synchronous architecture for simplicity and as all of our GPUs have roughly the same compute power.
+In this paper we work with a synchronous architecture for simplicity, as all of our GPUs have roughly the same compute power.
 There is also the argument of a centralised versus a decentralised topology.
@@
 
-We find in this paper \cite{lian2017can} that decentralized topology's are effected less by low bandwidths, and that is why we implement a decentralized architecture in our paper.
-To be more specific, we use our own implementation of the Baidu Ring AllReduce algorithm to average our gradients.
+We find in this paper (Lian, 2017)\cite{lian2017can} that decentralized topology's are effected less by low bandwidths, and that is why we implement a decentralized architecture in our paper.
+To be more specific, we use our own implementation of the Ring AllReduce algorithm \cite{gibiansky2017bringing} to average our gradients.
+
+All of our code will be written using the JAX \cite{jax2018github} machine learning framework with MPI being used for peer-to-peer communication via the mpi4jax \cite{mpi4jax} library.
 
@@
 Where $L$ is the loss function.
-We typically use an algorithm such as stochastic gradient descent(SGD) to solve the equation above.
+We typically use an algorithm such as Stochastic Gradient Descent(SGD) to solve the equation above.
 In which we need to calculate the unbiased stochastic gradients of $f$ at time step $t$ which satisfies $\mathbb{E}[g_t(w_{t-1})] = {\nabla} L(w_{t-1})$.
@@
 For example, suppose we have a set of 100 bins, each representing a range of gradient values.
-If the gradient of a particular parameter falls within the range of the first bin, it will be represented as a 1; if it falls within the range of the second bin, it will be represented as a 2; and so on.
+If the gradient of a particular parameter falls within the range of the first bin, it will be represented as a 0; if it falls within the range of the second bin, it will be represented as a 1; and so on.
 Then once the quantized $a$ has been sent to another node that node can look up what values the bin represents and replaces all the gradients in the bin with that value.
 This allows us to represent each gradient value using a single integer, rather than a floating-point value, which can significantly reduce the amount of bandwidth required to when sending gradients.
-The difference between using 8bit or 16bit integer is the number of bins, which effect not only the accuracy of the data but how much data is sent over the network.
+The difference between using 8bit or 16bit integer is the number of bins, which effects not only the accuracy of the data but how much data is sent over the network.
 If we use 16 bit integers there can be 65,536 bins where as with 8bit there is only 256 which is a significant reduction.
@@
 
-As mentioned above $a \sim \mathcal{N}(0.1,0.55)$ this tells us that that most of the values will be located around $0.1$ which means it may be more representative of $a$ if the there are more bins located around the mean than $\max a$ and $\min a$ so bins near the mean cover a lower range of values meaning they are occur at a higher frequency, bins closer to the min and the max cover a larger range of values.
+Our implementation of non-Linear quantization is a dumbed down version of \cite{dettmers20218}.
+As mentioned above $a \sim \mathcal{N}(0.1,0.55)$ this tells us that that most of the values will be located around $0.1$ which means it may be more representative of $a$ if the there are more bins located around the mean than $\max a$ and $\min a$.
+This means bins closer to the mean cover a lower range of values, but occur at a higher frequency.
+Bins closer to the min and the max cover a larger range of values.
 Instead of all the bins covering an equal range of values at a common interval each bin will contain an equal range of values(if you have 200 gradients and 10 bins each bin will contain 20 gradients).
@@
 If we represent all of those gradient as 256 different integers we loose a lot of accuracy and make it alot harder for the model to converge, so what we do is we split gradients into chunks of around 2000 or 4000 items(depending on the size of the model) and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk.
-there are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core makeing the quantization step a lot faster.
-If one of the values in $a$ was $7$ it is a clear outlier, this will effect one of the bins causing 1 bin to be wasted because of single outlier.
-By splitting up the data into chunks only a one of chunks would be effected by this outliers.
+There are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core makeing the quantization step a lot faster.
+If one of the values in $a$ was $7$ it is a clear outlier, this will effect one of the bins it to be wasted because of single outlier.
+By splitting up the data into chunks only a one chunk would be effected by this outliers.
 
 \subsection{Gradient Sparsification}
-Gradient sparsification is a technique that involves identifying and removing "zero" or nearly zero gradients from the gradients being transmitted between GPUs or computers in a distributed data parallelism setting.
+Gradient sparsification is a technique that involves identifying and removing "zero" or nearly zero gradients \cite{NEURIPS2018_3328bdf9} from the gradients being transmitted between GPUs or computers in a distributed data parallelism setting.
 The goal is to reduce the size of the gradients without significantly affecting their quality, so that they can be transmitted more efficiently over the network.
@@
 
-One negative effect of large batch sizes is that they can lead to a decrease in generalization performance.
-This occurs because large batch sizes can result in an increase in the variance of the stochastic gradient descent (SGD) updates, which can cause the model to overfit to the training data.
-This overfitting can manifest as a decrease in performance on held-out test data or a decrease in the ability of the model to generalize to unseen data.
+One negative effect of large batch sizes is that they can lead to a decrease in generalization performance \cite{keskar2016large}.
+This occurs because large batch sizes can result in an increase in the variance of the stochastic gradient descent (SGD) updates.
 
-Another negative effect of large batch sizes is that they can lead to a decrease in the convergence rate of the model.
+Another negative effect of large batch sizes is that they can lead to a decrease in the convergence rate of the model \cite{goyal2017accurate}.
 This occurs because large batch sizes can result in a reduction in the magnitude of the SGD updates, which can slow down the convergence of the model.
@@
 
-One way to counteract the second point is to increase the learning rate $\eta$ so that the magnitude of the SGD updates are increased this usually works for a few of epochs but after a while has a limited effect.
+One way to counteract the the decrease in convergence is to increase the learning rate $\eta$ \cite{goyal2017accurate} so that the magnitude of the SGD updates are increased this usually works for a few of epochs but after a while has a limited effect.
 We plan to use a adaptive batch size \cite{devarakonda2017adabatch}.
-At the start of training a an initial batch size is determined, and the primary test error is calculated from the models untrained parameters.
-After each epoch the the the test error is checked, once the test error is half of the primary test error the batch size gets doubled, once the test error is a quarter of the primary test error the batch size is doubled again batch.
+At the start of training an initial batch size is determined, and the primary test error is calculated from the models initial untrained parameters $w_0$.
+After each epoch the test error is checked, once the test error is half of the primary test error the batch size gets doubled, once the test error is a quarter of the primary test error the batch size is doubled again batch.
 
-Another way which we can reduce the negative effect of a large batch size is to use a different optimizer algorithm to SGD, a common alternative is Adam (Adaptive Moment Estimation) the problem with ADAM is that ADAM has been found to scale poorly with large batch sizes.
+Another way which we can reduce the negative effect of a large batch size is to use a different optimizer algorithm to SGD, a common alternative is Adam (Adaptive Moment Estimation) \cite{kingma2014adam} the problem with ADAM is that Adam has been found to scale poorly with large batch sizes.
 There are 2 optimizers that scale well with large batch sizes, LARS (Layer-wise Adaptive Rate Scaling) \cite{you2017scaling} \& LAMB (Layer-wise Adaptive Moments for Batch Training) \cite{you2019large}.
@@
 
-LAMB is pretty similar to ADAM.
-We start by computing the gradients $g_t$ at time step $t$, then much like ADAM we calculate the first $m$ and second $v$ moment of each layer $l$ using the following equations:
+LAMB is pretty similar to Adam.
+We start by computing the gradients $g_t$ at time step $t$, then much like Adam we calculate the first $m$ and second $v$ moment of each layer $l$ using the following equations:
 \begin{equation*} \label{1mom}
@@
 
-In these equations $\beta _{1/2}$ are the coefficient of the first/second moment. After we calculate the moments we trust ratio:
+In these equations $\beta _{1/2}$ are the coefficient of the first/second moment. After we bias correct the moments we calculate the trust ratio:
 
@@
 		r_1/r_2 & \quad r_1 > 0 \text{ and } r_2 > 0 \\
-		1 & \quad r_1 < 0 \text{ or } r_" < 0
+		1 & \quad r_1 < 0 \text{ or } r_2 < 0
 	\end{array}
@@
 \end{equation*}
-$r_1$ is the norm of the gradients and $r_2$ is the element-wise weight decay.
-To make sure that our value is $r^l_t$ isn’t too large or too small we have and upper and lower boundary $a$ and $b$.
-$\epsilon$ in equation \ref{r2} is to prevent a division by 0 error.
+$r_1$ is the norm of the gradients and $r_2$ is the element-wise weight decay. $r^l_t$ has to be greater than zero. $\epsilon$ in equation \ref{r2} is to prevent a division by 0 error.
 
@@
 
-These equation explain adaptive batch sizing.
+These equations explain adaptive batch sizing.
 First the scaling factor $\varepsilon$ where $E_0(w_0,x_R)$ is the primary test error, we make sure that value of $\varepsilon$ is greater than 1 as if for some reason $E_0(w_0,x_R) < E_e(w_t, x_R)$ the batch size would decrease which is not what we want.
 Then its just a case of multiplying $b$ by your scaling factor $\varepsilon$ and checking to see if its greater than $\beta$.
-Please not that using the output of a loss function in stead of the test error is sufficient.
+Please note that using the output of a loss function instead of the test error is sufficient.
 
@@
 The most important factor when choosing a model is to make sure we get a good range of sizes between our models.
-These are the models we train to get results:
+These are the models \cite{he2016deep,simonyan2014very,szegedy2017inception} we train to get results:
 
@@
 \subsection{Baseline}
-In out result we plan to compare the theoretical with the applied.In this case 8 GPUs were each node has 1 GPU, the data is evenly split between each node translating to each node training with a sample of 6250 images.
+In our result we plan to compare the theoretical with the applied.
+In this case 8 GPUs were each node has 1 GPU, the data is evenly split between each node translating to each node training with a sample of 6250 images.
 By increasing the number of nodes from 1 to 8 we reduces the number of communication steps per epoch to around 200 but there is also more data sent in each communication step as each node is sending its own gradients to 8 other nodes.
@@
 If we were dealing with our worst case scenario where you are training VGG16 on 100MBps you could take multiple hours ($4*200*56$ seconds) per epoch.
-This is cause by only 1 node sending data at a time while the rest of the nodes are idol caused by the networks constraints.
+This is caused by only 1 node sending data at a time while the rest of the nodes are idle caused by the networks constraints.
 As we are using a more efficient ring all reduce algorithm, we should have a much lower communication time than multiple hours.
@@
 \section{Conclusion}
-We were not able to access a computer so were not able to validate whether the techniques we implemented had a beneficial effect on the training time.
+We were not able to access a computer cluster, so were not able to validate whether the techniques we implemented had a beneficial effect on the training performance.
+We believe AI and Machine learning is getting to a point where it is completely inaccessible for the majority of researchers.
+Models are beginning to reach the 1 trillion parameter mark, making it impossible for a researcher that doesn't work for a company like Microsoft or Google to work on the cutting edge.
+We believe that by implementing techniques like the ones presented in this paper, machine learning at larger scale can become more accessible to everyone.
 
+\newpage
 \bibliographystyle{plain}

### 2023-02-20 paper/paper.tex
updates to format

@@
-\documentclass[11pt]{article}
+\documentclass[twocolumn,10pt]{article}
 \usepackage[margin=1in]{geometry}
@@
 \usepackage{amsmath, amsfonts,tabularx,graphicx, subfigure, mathtools}
+%\usepackage{ACL2023}
+\usepackage{multicol}
+\setlength{\columnsep}{1cm}
 \title{Improving the Performance of Distributed Data Parallelism \\Over Low Communication  Bandwidths}
@@
 \end{abstract}
-\tableofcontents
-\newpage
+%\tableofcontents
+%\newpage
+
+%\begin{multicols}{2}
 
@@
 
-Low WiFi bandwidths can have a significant impact on the training time of AI models using distributed data parallelism.
+Low communication bandwidths can have a significant impact on the training time of AI who are training using distributed data parallelism.
 When the communication channels between nodes have low bandwidth, this can lead to slower data transfer and synchronization between the nodes, which can in turn result in slower training times.
-In addition, low WiFi bandwidths may also lead to increased latency in the communication between nodes, further diminishing the efficiency of the training process.
+In addition, low communication bandwidths may also lead to increased latency in the communication between nodes, further diminishing the efficiency of the training process.
 
-In this paper, we aim to explore the relationship between low WiFi bandwidths and the training time of AI models using distributed data parallelism.
-We will discuss the various ways in which low WiFi bandwidths can impact the communication between nodes and the overall efficiency of the training process.
-We will also present potential solutions for mitigating the negative effects of low WiFi bandwidths on model training time.
+In this paper, we aim to explore the relationship between low communication bandwidths and the training time of AI models using distributed data parallelism.
+We will discuss the various ways in which low communication bandwidths can impact the communication between nodes and the overall efficiency of the training process.
+We will also present potential solutions for mitigating the negative effects of low communication bandwidths on model training time.
 
-To provide a better understanding of the role that WiFi bandwidth plays in the training of AI models using distributed data parallelism, we will conduct experiments using a variety of datasets and model architectures.
-We will compare the training time of models trained using distributed data parallelism in high and low WiFi bandwidth environments to evaluate the impact of low WiFi bandwidths on the training process.
+To provide a better understanding of the role that communication bandwidths play in the training of AI models using distributed data parallelism, we will conduct experiments using a variety of datasets and model architectures.
+We will compare the training time of models trained using distributed data parallelism in high and low communication bandwidth environments to evaluate the impact of low communication bandwidths on the training process.
 
-Overall, our goal is to shed light on the importance of WiFi bandwidth in the training of AI models using distributed data parallelism and to identify strategies for optimizing the training process in low bandwidth environments.
+Overall, our goal is to shed light on the importance of the communication bandwidth in the training of AI models using distributed data parallelism and to identify strategies for optimizing the training process in low bandwidth environments.
 We hope that the findings of this research will be useful for practitioners seeking to improve the efficiency of their model training process and for researchers studying the optimization of distributed data parallelism for AI model training.
@@
 On each GPU or computer in the cluster, the model 				  processes the data assigned to it and computes the gradients of the loss function with 		  respect to the model parameters.
-	\newpage
+	%\newpage
 	\item The gradients computed on each GPU or computer are then averaged together, and the 				  model parameters are updated using this average gradient.
@@
 It works by reducing the precision of the gradients that are transmitted between worker nodes.
-
+\\\\
 A model parameter is usually stored in memory as a 32bit floating point.
-There are $7, 800, 000, 000,\\ 000, 000,000,000,000,000,000,000,000,000,000,000,000,000,000$(38zeros) unique 32bit floating points, but that said 32bit floating points like $0.000100$, $0.000101$ and $0.000102$ are very close.
+There are $7.8\text{e}39$ (38zeros) unique 32bit floating points, but that said 32bit floating points like $0.000100$, $0.000101$ and $0.000102$ are very close.
 We can approximate these numbers to be equal, this is the core concept of quantization.
@@
 \end{equation*}
-$\phi(\parallel x \parallel)$ where $x \in \mathbb{R}^d$ is the same as doing $\sqrt{\sum^d_{i=1} x_i^2}$
+%$\phi(\parallel x \parallel)$ where $x \in \mathbb{R}^d$ is the same as doing $\sqrt{\sum^d_{i=1} x_i^2}$
 \begin{equation*}\label{r2}
@@
 \begin{center}
-\begin{tabular}{| p{3cm} | p{3cm} | p{3cm} | p{3cm} |} 
+\begin{tabular}{| p{2.2cm} | p{2.2cm} | p{2.2cm} |} 
 \hline
@@
 
-\newpage
 \section{Results}
@@
 
+\newpage
 \section{Conclusion}
@@
 \newpage
+%\newpage
 \bibliographystyle{plain}
@@
 
+%\end{multicols}
+
 \end{document}

### 2023-02-20 paper/paper.tex
updates

@@
 Modern models can contain billions of trainable parameters, which translates to billions of gradients.
-If we represent all of those gradient as 256 different integers we loose a lot of accuracy and make it alot harder for the model to converge, so what we do is we split gradients into chunks of around 2000 or 4000 items(depending on the size of the model) and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk.
+If we represent all of those gradients as 256 different integers we loose a lot of accuracy and make it alot harder for the model to converge, so what we do is we split gradients into chunks of around 2000 or 4000 items(depending on the size of the model) and quantize each chunk separately this means that the first bin in the first chunk might cover a different ranges of values to the first bin in the second chunk.
 There are a few reasons for this, the first being that it allows us to use 8bit integers without a significant loss in accuracy, the second being that we can allocate each chunk to a specific GPU core makeing the quantization step a lot faster.
-If one of the values in $a$ was $7$ it is a clear outlier, this will effect one of the bins it to be wasted because of single outlier.
+If one of the values in $a$ was $7$ it is a clear outlier, this will effect one of the bins which means it would be wasted because of single outlier.
 By splitting up the data into chunks only a one chunk would be effected by this outliers.
@@
 
-In order to perform gradient sparsification we have to decide on a threshold to determine whether or not a gradient is a zero gradient or not, if the gradient is less than the threshold then it it deemed a zero gradient.
+In order to perform gradient sparsification we have to decide on a threshold to determine whether or not a gradient is a zero gradient.
+If the gradient is less than the threshold then it is deemed a zero gradient.
 In our implementation of gradient sparsification we accumulate zero gradients locally and sum them with gradients from the next iteration.
